---
title: "Lab 18: Inference for the Mean"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Teacher Salary

We will start today by looking at a small dataset containing teacher
salaries from 2009-2010 for 71 randomly choosen teachers employed by the
St. Louis Public School in in Michigan.

```{r}
teachers <- read_csv("https://statsmaths.github.io/stat_data/teachers_pay.csv")
```

The available variables are

- base: Base annual salary, in dollars
- degree: Highest educational degree attained: BA (Bachelor's) or MA (Master's)
- years: Number of years employeed

Using the mean function, what is the average base pay of all teachers
in the dataset?

```{r}
mean(teachers$base)
```

Fit a model for the mean of the base pay variable using lm_basic. Save
the model as an object called "model":

```{r}
model <-lm_basic(base ~ 1, data = teachers)
```

Using a call to `reg_table`, find the mean implied by the model:

```{r}
reg_table(model)
```

Does the mean agree with your answer to question 2?

**Answer**: Yes

Add a 95% confidence interval to the regression table.

```{r}
reg_table(model, level = 0.95)
```

What is the range of mean salaries implied by the confidence interval?

**Answer**: The range of mean salaries by the confidence interval is 56938 to 58737

Draw a histogram of the base salary values for the entire dataset.

```{r}
hist(teachers$base)
```

Do most of the salary values fall within the range given in question 5? Why
or why not?

**Answer**: Yes, most of them fall within the similar mid 50000 to upper 50000 range. 

Use the filter command to construct a new dataset called `masters` consisting
of just those teachers with a masters degree.

```{r}
masters <- filter(teachers, degree == "MA")
```

Compute a 95% confidence interval for the mean pay of teachers with a master's
degree. Does this range intersect the one you had in question 5?

```{r}
model2 <-lm_basic(base ~ 1, data = masters)

reg_table(model2, level = 0.95)


```
This range is higher than the previous base range by about 2000. 

## Murder Data

Now load the following dataset containing all murders that have occurred in
London from 1 January 2006 to 7 September 7 2011.

```{r}
london <- read_csv("https://statsmaths.github.io/stat_data/london_murders.csv")
```

The available variables are:

- age: age of the victim in years
- year: year of the murder
- borough: the London borough in which the murder took place

Find an 80% confidence interval for the average age of the victim of a
murder in London.

```{r}
modellondon<- lm_basic(age ~ 1, data = london)

reg_table(modellondon, level = .80)
```

Make sure you actually extract the answer here:

**Answer**: The average age range of murder victims in London are between 33 and 34. 

Describe in words how the proceeding confidence interval should be
interpreted.

**Answer**: 80% of the time, the murder victim in London will be in their early to mid 30s. However, there is a decent standard error that it may fall outside this range. If we had a 90% or 95% confidence interval, we would be able to be more narrow and precise in our range, but for 80% most of the people are betwen 33 and 34. 








