---
title: "Lab 17: Numeric Summaries"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Cereal Data

Today, we start by looking at a collection of breakfast cereals:

```{r}
cereal <- read_csv("https://statsmaths.github.io/stat_data/cereal.csv")
```

With variables:

- name: name of the specific cereal
- brand: name of the cereal's manufacturer
- sugar: amount of sugar per serving (g)
- score: healthiness score; 0-100; 100 is the best
- shelf: what shelf the cereal is typically stocked on in the store

Produce a histogram of the sugar variable.

```{r}
hist(cereal$sugar)
```

Now, compute the standard deviation of the variable `sugar`:

```{r}
sd(cereal$sugar)
```

What are the units of this measurement?


**Answer**: The units of measurement for sugar is amoung of sugar per serving in grams. 

Now, compute the deciles of the variable `score`:

```{r}
deciles(cereal$score)
```

What is the value of the 30th percentile. Describe what this means in words:

**Answer**: 34.5 is the 30%. This means that a third of cereal have a healthiness score of 34.5, which is very low considering 100 equals the healthiest. 

Produce a boxplot of score and brand.

```{r}
ggplot(cereal, aes(brand, score))+
  geom_boxplot()
```

Which brand seems to have the healthiest cereals?

**Answer**: Nabisco seems to have the healthiest cereals with their average score being above 60. All of the other cereals have a large standard devidation averaging around the 40s. 

Produce a boxplot of score and shelf.

```{r}
ggplot(cereal, aes(shelf, score))+
  geom_boxplot()
```

Produce a boxplot of sugar and shelf.

```{r}
ggplot(cereal, aes(shelf, sugar))+
  geom_boxplot()
```

If I want a healthy but reasonably sweet cereal which shelf would be the
best to look on?

**Answer**: The shelf that has the healthiest, but reasonable sweet cereal is the bottom shelf, as it has a lower sugar quantity. However, the top shelf also has cereals that fit this category. Only the middle shelf has an obvious high sugar rating with a low healthiness rating. 

## Tea Reviews

Next, we will take another look at a dataset of tea reviews that I used in
a previous lecture:

```{r}
tea <- read_csv("https://statsmaths.github.io/stat_data/tea.csv")
```

With variables:
- name: the full name of the tea
- type: the type of tea. One of:
    - black
    - chai
    - decaf
    - flavors
    - green
    - herbal
    - masters
    - matcha
    - oolong
    - pu_erh
    - rooibos
    - white
- score: user rated score; from 0 to 100
- price: estimated price of one cup of tea
- num_reviews: total number of online reviews

Draw a scatterplot with num_reviews (x-axis) against score (y-axis) and add a
regression line (recall: `geom_smooth(method="lm")`).

```{r}
ggplot(tea, aes(num_reviews, score))+
  geom_point()+
  geom_smooth(method="lm")
```

Does the score tend to increase, decrease, or remain the same as the number
of reviews increases?

**Answer**: The score tends to increase as the number of reviews increase. 

Calculate the ventiles of the variable price.

```{r}
ventiles(tea$price)
sd(tea$price)
```

What is the 80th percentile? Describe it in words, include the units of the
problem in your answer.

**Answer**: For 80% of the teas, the price of a cup of tea is estimated to be $30 dollars. 

Plot the number of reviews (x-axis) against the score variable. Color
the points according to price binned into 5 buckets.

```{r}
quartiles(tea$price)
ggplot(tea, aes(price, num_reviews)) +
  geom_point(aes(color = bin(price, 5)))
```

What tends to be true about the number of reviews for the most expensive
20% of teas?

**Answer**: For the most expensive 20% of teas, there are less reviews because who would pay over $30 for a cup of tea??? In seriousness, most people would not pay for that price, so there is ultimately less reviews. 

Create a dataset named `white` that consists of only white teas.

```{r}
whites <- filter(tea, type == "white")
whites
```

Calculate the standard deviation of the price for white teas and the
standard deviation of the price for all of the teas.

```{r}
sd(whites$price)
```

Is the variation of the white tea prices smaller, larger, or about the same
as the entire dataset?

**Answer**: The variation of the white tea prices are smaller than the entire dataset. The white tea standard deviation is 13.59 while overall was 30.42.

Summarize the dataset by the type of tea and save the results as a variable
named `tea_type`.

```{r}
tea_type <- group_summarize(tea, type)
```

Plot the average price (x-axis) against the average score (y-axis) of
each type  of tea. Make the size of the points proportional to the number
of teas in each category and label the points with geom_text_repel and the
tea type.

```{r}
library(ggrepel)
ggplot(tea_type, aes(price_mean, score_mean))+
  geom_text_repel(aes(price_mean, score_mean, label = rownames(type)))+
  theme_classic(base_size = 10)
```

Describe an interesting pattern or set of outliers that you found in the
previous plot. This does not need to take more than 1-2 sentences.

**Answer**:


