---
title: "Chilling in Colorado"
author: "Byongho Lee"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
```


```{r, include=FALSE}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
```


```{r, include=FALSE}
Denver<- filter(tracts, state == "Colorado")
```

```{r, include=FALSE}
library(ggmap)
```

```{r, include=FALSE}
acs_den <- filter(tracts, cbsa == "Denver-Aurora-Lakewood, CO")
wealth_area <- filter(acs_den, hh_income_top_5>77000)
```

## Denver, Colorado 

I chose Denver, Colorado beause I lived there for three months for a summer internship in 2017. I remember having trouble finding a place to stay that was both affordable and close to my internship. My internship location was in the center of downtown, very close to Rockies Stadium. I ended up living in Highland Park, about forty minutes away walking distance and twenty by public transportation. My goal is to uncover interesting information about living in the city of Denver and use my own experiences and profile to find information about the general demographics.
```{r, include=FALSE}
qmplot(lon, lat, data = acs_den, geom = "blank")
```

### Hard to Find Housing?

First, I will be looking at specific demographics such as race and income equality to see their relationship with rent costs. In cities, rent tends to be higher than rural areas. For places such as New York or Los Angeles, college grauates and people with limited wealth migrate to these areas due to better opportunities. However, Denver has reasonable rent prices with the median rent being between 1000 to 2000. The graph below shows the conentration of rentable living areas in the center of Denver. The interesting element in this graph is that the farther out the location is, the more expensive the median rent is. When I was trying to find housing, it seemed most places cost around 1500-2200. This number is high for a college student on an internship salary, but reasonable for a full-time worker. I might have found a cheaper apartment if I looked closer in the city. 

```{r}
qmplot(lon, lat, data = acs_den, geom = "blank") +
  geom_point(aes(color = median_rent), alpha = 0.8) +
  scale_color_viridis(option="inferno")+
  labs(color = "Median Rent") +
  ggtitle("Rent Costs in Denver")
```

### Who Looks Like Me?

The second variable I wanted to observe was the racial demographics of Denver. When I was in Denver, I noticed there was not a lot of diversity in the city. I would take weekend trips with some local friends and I saw more blacks, asians, and hispanics outside the city. The graph shoes the strong concentration of whites in Denver. This has implications for Colorado as a whole; cities tend to have certain areas dedicated towards different groups such as Chinatowns and Koreatowns. Denver is growing and we may see more diversity in the next few years. Colorado needs to make sure that it doesn't intentionally segregate different groups although it may end up happening that way.   


```{r}
qmplot(lon, lat, data = acs_den, geom = "blank") +
  geom_point(aes(color = race_white), alpha = 0.8) +
  scale_color_viridis(option = "plasma")+
  labs(color = "Perentage of Whites") +
  ggtitle("White People In Denver")
```
Exploring the racial demographic further, we use the same process and data to find the concentration of asians in Denver. Similarily, it shows low rates for minorities in the downtown and Lakewood areas. This graph is consistent with the one above with whites being the dominating population. What is interesting is that Asians are not concentrated in the central Aurora area, which is where there is a low concentration of whites. Asians centralize in the Centenniel area or near Cherry State Park, indicated by the orannge, yellow points. Overall, Denver is a predominantly white area and I noticed that during my summer internship. 

```{r}
qmplot(lon, lat, data = acs_den, geom = "blank") +
  geom_point(aes(color = race_asian), alpha = 0.8) +
  scale_color_viridis(option = "plasma")+
  labs(color = "Percentage of Asians") +
  ggtitle("Asian People In Denver")
```

Finally, the last graph tries to analyze the concentration of wealth for individuals in Denver. For the most part, Denver has a gini coefficient of .3, which shows a decent amount of income equality. The highest amount of income inequality is near the center of Denver with coefficients of upper .5s and .6s. Comparing this graph to previous graphs, we can see that while income is more equal outside the city, rent costs are higher. It may be that people of normal to higher incomes concentrate outside the city. In relation to the previous graphs, most of the white demographic have higher rates of income inequality in the city. 

```{r}
qmplot(lon, lat, data = wealth_area, geom = "blank") +
  geom_point(aes(color=hh_gini), alpha = 0.8) +
  scale_color_viridis(option = "cividis")+
  labs(color = "Gini Coefficient") +
  ggtitle("Income Equality/Inequality")
```

Overall, the average rent costs in Denver are relatively low. As an asian male with a student budget, the cost of living was manageable for a summer. Using the data gathered, I was able to create four graphs that depicted the relationship between rent costs, race, and income inequality. Ultimately, Denver is a predominately white area that has low rent costs in the city and low to medium income inequality. 


