---
title: "Chilling in Colorado"
author: "Byongho Lee"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
```


```{r, include=FALSE}
tracts <- read_csv("https://raw.githubusercontent.com/statsmaths/stat_data/gh-pages/acs_tract_data_2016.csv.gz")
```


```{r, include=FALSE}
Denver<- filter(tracts, state == "Colorado")
```

```{r, include=FALSE}
library(ggmap)
```

```{r, include=FALSE}
acs_den <- filter(tracts, cbsa == "Denver-Aurora-Lakewood, CO")
wealth_area <- filter(acs_den, hh_income_top_5>77000)
```

## Denver, Colorado 

I chose Denver, Colorado beause I lived there for three months for a summer internship in 2017. I remember having trouble finding a place to stay that was both affordable and close to my internship. My internship location was in the center of downtown, very close to Rockies Stadium. I ended up living in Highland Park, about forty minutes away walking distance and twenty by public transportation. My goal is to uncover interesting information about living in the city of Denver and use my own experiences and profile to find information about the general demographics.
```{r, include=FALSE}
qmplot(lon, lat, data = acs_den, geom = "blank")
```

### Hard to Find Housing?

First, I will be looking at specific demographics such as race and income equality to see their relationship with rent costs. In cities, rent tends to be higher than rural areas. For places such as New York or Los Angeles, college grauates and people with limited wealth migrate to these areas due to better opportunities. However, Denver has reasonable rent prices with the median rent being between 1000 to 2000. The graph below shows the conentration of rentable living areas in the center of Denver. The interesting element in this graph is that the farther out the location is, the more expensive the median rent is. When I was trying to find housing, it seemed most places cost around 1500-2200. This number is high for a college student on an internship salary, but reasonable for a full-time worker. I might have found a cheaper apartment if I looked closer in the city. 

```{r}
qmplot(lon, lat, data = acs_den, geom = "blank") +
  geom_point(aes(color = median_rent), alpha = 0.8) +
  scale_color_viridis(option="inferno")+
  labs(color = "Median Rent") +
  ggtitle("Rent Costs in Denver") +
  annotate("point", x = -105.0003, y = 39.842, color="red", size=5) +
  annotate("text", x= -105.423, y = 39.90, label = "Highland Park")
```

### Who Looks Like Me?

The second variable I wanted to observe was the racial demographics of Denver. When I was in Denver, I noticed there was not a lot of diversity in the city. I would take weekend trips with some local friends and I saw more blacks, asians, and hispanics outside the city. The graph below shows the percentage of white people in areas surrounding Denver. I use age and race to see the percentage of diversity in the Denver areas and it seems most areas are dominated by the white population. This has implications for Colorado as a whole; major U.S. cities tend to have certain areas dedicated towards different groups such as Chinatowns and Koreatowns. Denver is growing, but for the most part, the white population is huge with most areas being above 75% white. 


```{r}
ggplot(acs_den, aes(race_white, age_18_24))+
  geom_point()+
  geom_point(aes(race_asian), color="blue")+
  xlab("Percentage of Race") +
  ylab("Percentage of People between ages 18-24") +
  ggtitle("White and Asians")+
  annotate("text", x=35, y=30, color="blue", label = "Blue means percentage of Asians")+
  annotate("text", x=50, y=35, label = "Black means percentage of Whites")
  
```
Exploring the racial demographic further, we use the same process and data to find the concentration of asians in Denver. The blue points indicate the percentage of asian populations in the Denver areas. Most areas do not exceed 25%, which means there is a small population of asians across every Denver area. When I visited Colorado, the asians were centralized in the Centenniel area or near Cherry State Park. In that area, we see that Asians have the highest percentage (25%). Overall, Denver is a predominantly white area and I noticed that during my summer internship. 

## Equal or Unequal?

Finally, the last graph tries to analyze the concentration of wealth for individuals in Denver. For the most part, Denver has a gini coefficient of .3, which shows a decent amount of income equality. The highest amount of income inequality is near the center of Denver with coefficients of upper .5s and .6s. Comparing this graph to previous graphs, we can see that while income is more equal outside the city, rent costs are higher. It may be that people of normal to higher incomes concentrate outside the city. In relation to the previous graphs, most of the white demographic have higher rates of income inequality in the city. 

```{r}
qmplot(lon, lat, data = wealth_area, geom = "blank")+
  geom_point(aes(color=hh_gini), alpha = 0.8)+
  scale_color_viridis(option = "cividis")+
  labs(color = "Gini Coefficient")+
  ggtitle("Income Equality/Inequality")
```
## Summary 

Overall, the average rent costs in Denver are relatively low. As an asian male with a student budget, the cost of living was manageable for a summer. Using the data gathered, I was able to create four graphs that depicted the relationship between rent costs, race, and income inequality. Ultimately, Denver is a predominately white area that has low rent costs in the city and low to medium income inequality. There are not many asians in the Denver area compared to other cities such as New York or Los Angeles. 


